A higher order approximate static condensation method for multi-material diffusion problems

### 2019-01-26 num_sec.tex
Add Maxim's changes

\section{Numerica results}\label{sec:num}

Let $h$ be a max cell diameter for the macro-mesh~$\bmesh_h$, and $\mathbb V_h \subset \LTwoSpace$ be a space of piecewise constant
functions on each cell~$\mcell \in \mmesh(\bcell)$, $\bcell \in \bmesh_h$. We define \textit{discrete} $\LTwo$-norm of $v \in \LTwoSpace$ as
\begin{equation}\label{l2}
  \| v \|_{\lTwoSpace} \coloneqq \| P_h\,v \|_{\LTwoSpace},
\end{equation}
where $P_h\;:\;\LTwoSpace\rightarrow\mathbb V_h$ is $\LTwo$-projection operator.

For the error~$e_h \coloneqq p - p_h$ between the exact and computed solutions we have
\begin{equation}\label{l2norm}
	\errlTwo{h} \coloneqq \| e_h \|_{\lTwoSpace} = \| p - p_h \|_{\lTwoSpace} = \| P_h\,p - P_h\,p_h \|_{\LTwoSpace} = \| P_h\,p - p_h \|_{\LTwoSpace}
\end{equation}
since~$P_h$ is linear and~$p_h \in \mathbb V_h$. The discrete norm~\eqref{l2norm} of the concentration discretization error is computed as follows:
$$
	\errlTwo{h} = \left[\sum\limits_{\bcell \in \bmesh_h}\sum\limits_{\mcell\in\mmesh(\bcell)} \left(\frac{1}{|\mcell|}\int_{\mcell} p \diff{\vect x} - p_h(\mcell)\right)^2|\mcell|\right]^\frac12,
$$
        where $|\mcell|$ is the area of cell $\mcell$. While the disrete norm of the flux discretization error is defined as follows:
        $$
        \varepsilon_h^{l_2} =  \left[\sum\limits_{\bface \in \bmesh_h}
          A_\bface\left(
          \sum\limits_{\mface\in\bface}
          \frac{1}{|\mface|}\int_{\mface}
          \vect u \diff{\vect s}  - \sum\limits_{\mface\in\bface}\vect u_h(\mface)\right)^2 
          \right]^\frac12
        $$
        where $A_\bface|$ is the representative are of a face $\bface$ which is equal to the arithmetic average
of areas of cells sharing the face $\bface$.

If~$\bmesh_h$ consists of triangles and no material interfaces are present, ASC($n$) boils down to mixed-hybrid Raviart\,--\,Thomas finite element method (which is algebraically equivalent to~$R\,T_0$ finite element method). In this particular case we have~\cite{M2AN_1980__14_3_249_0}
$$
	\|\vect u - \vect u_h\|_{\LTwoSpace} \le c\,h\,\|\vect u\|_{\HSpace{1}},\quad\|p - p_h\|_{\LTwoSpace} \le c \left( h\,\|p\|_{\HSpace{1}} + h^2\,\|p\|_{\HSpace{2}} \right),
$$
so one cannot expect ASC($n$) $\LTwo$-convergence to be better than linear. Note that
\begin{equation}\label{l2L2}
	\| p - p_h \|_{\lTwoSpace} \le \| p - p_h \|_{\LTwoSpace}
\end{equation}
since $\| p - p_h \|_{\lTwoSpace} = \| P_h\,(p - p_h) \|_{\lTwoSpace} \le \|P_h\|_{\mathcal L(\LTwoSpace, \mathbb V_h)} \| p - p_h \|_{\LTwoSpace}$, and the operator norm~$\|P_h\|_{\mathcal L(\LTwoSpace, \mathbb V_h)} = 1$ by Pythagorean theorem. Thus one may expect some improvement in $\lTwo$-convergence as one increases~$n$ for ASC($n$).

@@

The first set of tests is designed to check the property of the method to accurately recover the solution that is the polynomial of degree~1. Let us consider the diffusion problem~\eqref{dual} with~$c \equiv f \equiv 0$, $\partial\Omega_D = \partial\Omega$. The computational domain $\Omega=(0,1)^2$ is divided into several subdomains by non-intersecting straight lines. In these settings the interface reconstruction algorithm MOF reconstructs the interfaces exactly. To test the {\it linearity preservation} property, we set up a pseudo-multi-material problem with the diffusion tensor being the same in all subdomains, $\vect K_i = \vect I$. The exact solution is a linear function. The geometry of two cases under consideration is shown in Figure~\ref{fig:fake}.

@@

Since ASC(1) approximates interface traces with $P_1$ functions it recovers edge-based degrees of freedom exactly in the sense of mean values. This results in exact reconstruction of cell-based unknowns and {\it linearity preservation} property for both examples, i.e.
\begin{equation*}
	\errlTwo{h,\,1} = \left\| p - p_{h,\,1} \right\|_{\lTwoSpace} = 0.
\end{equation*}
% As we saw in previous section,
%In general, one may not expect concentration and flux convergence to be better than second order with respect to~$\lTwo$\,--\,error for ASC($1$).  However, taking the example above into account, one may expect that convergence of ASC(1) with respect to~$\lTwo$-norm is superior
%to  convergence properties of ASC(0).

@@

For this example we use a sequence of square base\,meshes~$\bmesh_{h_i}$, $i = 1, 2, 3, 4$. We denote by
\begin{equation*}
	\rho_{h_i,\,n} \coloneqq \frac{\ln \errlTwo{h_i,\,n} / \errlTwo{h_{i+1},\,n}}{\ln h_i / h_{i+1}}
\end{equation*}
the reduction rate order of ASC($n$) in~$\lTwo$-norm between $i$th and $(i+1)$th refinement steps. We also compute the $\ell^{\infty}$-norm of the error
\begin{equation*}
	\errInf{h_i,\,n} \coloneqq \left\| p - p_{h_i,\,n} \right\|_\infty
\end{equation*}
for ASC($n$).

@@

Numerical results are shown in Table~\ref{fig:conv:pwlin}. We observe that ASC(0) converges linearly with respect to the $\lTwo$-norm,
and ASC(1) has the convergence rate close to quadratic with respect to the $\lTwo$-norm.
The convergence of the flux errors is also is higher for ASC(1) method and is close to linear convergence.

@@

It is well-known that the presence of a few outliers in the spectrum does not affect the asymptotic convergence of the conjugate gradient (CG) iterative methods, see, e.g., \cite{olshanskii2014iterative}. Indeed, in our experiments the CG method (with algebraic multigrid preconditioner) was found to be equally effective for solving systems of algebraic equations resulting from ASC(0) and ASC(1).
%Starting from some iteration CG method behaves like extreme eigenvalues are not present in the spectrum.
%That is, several small eigenvalues with no clustering is not an obstacle neither for the optimal convergence rate of Krylov solvers, nor for the~$\LTwo$-convergence of the discrete solution.

@@

The computational domain is the same as in section~\ref{sec:threemat} with the example of three different materials.  $\Omega_1 = \{{\bf x} : \|{\bf x} - {\bf x}_0\| < 0.15\}$, $\Omega_2 = \{{\bf x} : 0.15 < \|{\bf x} - {\bf x}_0\| < 0.2\}$, and $\Omega_3 = (0,1)^2 \setminus (\bar{\Omega}_1\cup\bar{\Omega}_2)$. The diffusion tensor is set as $\vect K_i = k_i\,\vect I$, $k_1 = 10, k_2 = 0.002$, and $k_3 = 1$.

@@

We first compute solution $p_*$ to the problem~(\ref{dual_transient}) using $P2$ finite element method on sufficiently fine mesh consisting of 4\,908 triangles. This mesh is illustrated in Figure~\ref{fig:transient:ref} (left). The reference mesh is consistent with material interfaces so it consists only of single-material cells. Time step~$\Delta t$ is chosen as~$T / 40$. This solution was found to be (almost) mesh independent and will serve as the  reference solution. The center cutline profiles~$p_*\big((x,0.5), t\big)$   are shown in the Figure~\ref{fig:transient:ref} (right) for several values of~$t \in (0, T]$.

For ASC(0), ASC(1), arithmetic and harmonic homogenization methods we use two Voronoi meshes with~$h = 1.5 \times 10^{-1}$ (124 base\,cells) and $h = 8.1 \times 10^{-2}$ (465 base\,cells).
The coarse mesh contains 20 multi-material cells and two macro-faces sharing three materials. The fine mesh contains 50 multi-material cells and no three material macro-faces. For all these methods we use 20 time frames, $\Delta t = T / 20$. Results are shown in Figures~\ref{fig:transient:comp}\,--\,\ref{fig:transient:err}.

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2/movie_frames/ref/frame_1s34.png}
		\caption{Reference}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/arithmetic_homo/frame_1s43.png}
		\caption{Arithmetic homo.}	
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/harmonic_homo/frame_1s43.png}
		\caption{Harmonic homo.}	
	\end{subfigure}%
	\par
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/asc0/frame_1s43.png}
		\caption{ASC(0)}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/asc1/frame_1s43.png}
		\caption{ASC(1)}	
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/asc_diff/frame_1s43.png}
		\caption{ASC(0,\,1) difference}	
	\end{subfigure}%
	\caption{Comparison of the discrete solutions~$p_h$ for~$h = 1.5 \times 10^{-1}$, $t = 1.25$ \label{fig:transient:comp}}
 	\centering
 	\vskip 1cm
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2-1/movie_slices_frames/frame_0s25.png}
 		\caption{Coarse mesh, $t = 0.25$}
 	\end{subfigure}%
 	\hfill
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2-1/movie_slices_frames/frame_1s50.png}
 		\caption{Coarse mesh, $t = 1.50$}	
 	\end{subfigure}%
 	\par
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2/movie_slices_frames/frame_0s25.png}
 		\caption{Fine mesh, $t = 0.25$}
 	\end{subfigure}%
 	\hfill
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2/movie_slices_frames/frame_1s50.png}
		\caption{Fine mesh, $t = 1.50$}	
	\end{subfigure}%
	\caption{Comparison of the cuts~$p_h\big((x,0.5), t\big)$ of discrete solutions for coarse and fine meshes, $h = 1.5 \times 10^{-1}$ and $h = 8.1 \times 10^{-2}$ 		\label{fig:transient:cuts}}
\end{figure}

\clearpage

>>>>

\section{Numerical results}\label{sec:num}

Let $h$ be a maximum cell diameter for the macro-mesh~$\bmesh_h$, and $\mathbb V_h \subset \LTwoSpace$ be a space of piecewise constant
functions on each cell~$\mcell \in \mmesh(\bcell)$, $\bcell \in \bmesh_h$. We define \textit{discrete} $\LTwo$-norm of $v \in \LTwoSpace$ as
\begin{equation}\label{l2}
  \| v \|_{\lTwoSpace} \coloneqq \| P_h\,v \|_{\LTwoSpace},
\end{equation}
where $P_h\;:\;\LTwoSpace\rightarrow\mathbb V_h$ is $\LTwo$-projection operator.

For the error~$e_h \coloneqq p - p_h$ between the exact and computed solutions we have
\begin{equation}\label{l2norm}
	\errlTwo{h} \coloneqq \| e_h \|_{\lTwoSpace} = \| p - p_h \|_{\lTwoSpace} = \| P_h\,p - P_h\,p_h \|_{\LTwoSpace} = \| P_h\,p - p_h \|_{\LTwoSpace}
\end{equation}
since~$P_h$ is linear and~$p_h \in \mathbb V_h$. %The discrete norm~\eqref{l2norm} of the concentration discretization error is computed as follows:
This gives
$$
	\errlTwo{h} = \left[\sum\limits_{\bcell \in \bmesh_h}\sum\limits_{\mcell\in\mmesh(\bcell)} \left(\frac{1}{|\mcell|}\int_{\mcell} p \diff{\vect x} - p_h(\mcell)\right)^2|\mcell|\right]^\frac12,
$$
        where $|\mcell|$ is the area of cell $\mcell$. The discrete norm of the flux discretization error is defined as follows:
        $$
        \varepsilon_h^{l_2} =  \left[\sum\limits_{\bface \in \bmesh_h}
          A_\bface\left(
          \sum\limits_{\mface\in\bface}
          \frac{1}{|\mface|}\int_{\mface}
          \vect u \diff{\vect s}  - \sum\limits_{\mface\in\bface}\vect u_h(\mface)\right)^2
          \right]^\frac12
        $$
        where $A_\bface$ is the reference volume of a face $\bface$ which is equal to the arithmetic average
of volumes of cells sharing the face $\bface$.

If~$\bmesh_h$ consists of triangles and \emph{no} material interfaces are present, ASC($n$) reduces to mixed-hybrid Raviart\,--\,Thomas finite element method (which is algebraically equivalent to~$R\,T_0$ finite element method). In this particular case we have~\cite{M2AN_1980__14_3_249_0}
$$
	\|\vect u - \vect u_h\|_{\LTwoSpace} \le c\,h\,\|\vect u\|_{\HSpace{1}},\quad\|p - p_h\|_{\LTwoSpace} \le c \left( h\,\|p\|_{\HSpace{1}} + h^2\,\|p\|_{\HSpace{2}} \right),
$$
so one should not expect ASC($n$) $\LTwo$-convergence to be better than linear. 
However, since we measure convergence of the averages values and
%\begin{equation}\label{l2L2}
$	\| p - p_h \|_{\lTwoSpace} \le \| p - p_h \|_{\LTwoSpace} $
%\end{equation}
by the property of projection, we may observe a higher order convergence for smooth solutions.
%since $\| p - p_h \|_{\lTwoSpace} = \| P_h\,(p - p_h) \|_{\lTwoSpace} \le \|P_h\|_{\mathcal L(\LTwoSpace, \mathbb V_h)} \| p - p_h \|_{\LTwoSpace}$, and the operator norm~$\|P_h\|_{\mathcal L(\LTwoSpace, \mathbb V_h)} = 1$ by Pythagorean theorem. Thus one may expect some improvement in $\lTwo$-convergence as one increases~$n$ for ASC($n$).

@@

The first set of tests is designed to check the property of the method to accurately recover the solution that given by a  polynomial of degree~1. Let us consider the diffusion problem~\eqref{dual} with~$c \equiv f \equiv 0$, $\partial\Omega_D = \partial\Omega$. The computational domain $\Omega=(0,1)^2$ is divided into several subdomains by non-intersecting straight lines. In these settings the interface reconstruction algorithm MOF reconstructs the interfaces exactly. To test this preservation property, we set up a pseudo-multi-material problem with the diffusion tensor being the same in all subdomains, $\vect K_i = \vect I$. The exact solution is a linear function. The geometry of two cases under consideration is shown in Figure~\ref{fig:fake}.

@@

Since ASC(1) approximates interface traces with $P_1$ functions it recovers edge-based degrees of freedom exactly in the sense of mean values. This results in exact reconstruction of cell-based unknowns and the preservation property holds for both examples, i.e.
\begin{equation*}
	\errlTwo{h,\,1} = \left\| p - p_{h,\,1} \right\|_{\lTwoSpace} = 0.
\end{equation*}
% As we saw in previous section,
%In general, one may not expect concentration and flux convergence to be better than second order with respect to~$\lTwo$\,--\,error for ASC($1$).  However, taking the example above into account, one may expect that convergence of ASC(1) with respect to~$\lTwo$-norm is superior
%to  convergence properties of ASC(0).

@@

For this example we use a sequence of square base\,meshes~$\bmesh_{h_i}$, $i = 1, 2, 3, 4$. We denote by
\begin{equation*}
	\rho_{h_i,\,n} \coloneqq \frac{\ln (\errlTwo{h_i,\,n} / \errlTwo{h_{i+1},\,n})}{\ln (h_i / h_{i+1})}
\end{equation*}
the convergence order of ASC($n$) in~$\lTwo$-norm between $i$th and $(i+1)$th refinement steps. We also compute the $\ell^{\infty}$-norm of the error
\begin{equation*}
	\errInf{h_i,\,n} \coloneqq \left\| p - p_{h_i,\,n} \right\|_\infty
\end{equation*}
for ASC($n$).

@@

Numerical results are shown in Table~\ref{fig:conv:pwlin}. We see that ASC(0) converges linearly with respect to the $\lTwo$-norm,
and ASC(1) has the convergence rate close to quadratic with respect to the same norm.
The convergence of the flux errors is also higher for ASC(1) method and is close to linear convergence.

@@

It is well-known that the presence of a few outliers in the spectrum does not affect the \emph{asymptotic} convergence of the conjugate gradient (CG) iterative methods, see, e.g., \cite{olshanskii2014iterative}. Indeed, in our experiments the CG method (with algebraic multigrid preconditioner) was found to be equally effective for solving systems of algebraic equations resulting from ASC(0) and ASC(1).
%Starting from some iteration CG method behaves like extreme eigenvalues are not present in the spectrum.
%That is, several small eigenvalues with no clustering is not an obstacle neither for the optimal convergence rate of Krylov solvers, nor for the~$\LTwo$-convergence of the discrete solution.

@@

The computational domain is the same as in section~\ref{sec:threemat} with the example of three different materials.  $\Omega_1 = \{{\bf x} : \|{\bf x} - {\bf x}_0\| < 0.15\}$, $\Omega_2 = \{{\bf x} : 0.15 < \|{\bf x} - {\bf x}_0\| < 0.2\}$, and $\Omega_3 = (0,1)^2 \setminus (\bar{\Omega}_1\cup\bar{\Omega}_2)$. The diffusion tensor is set to be $\vect K_i = k_i\,\vect I$, $k_1 = 10, k_2 = 0.002$, and $k_3 = 1$.

@@

We first compute solution $p_*$ to the problem~(\ref{dual_transient}) using $P_2$ finite element method on sufficiently fine mesh consisting of 4\,908 triangles. This mesh is illustrated in Figure~\ref{fig:transient:ref} (left). The reference mesh is consistent with material interfaces so it consists only of single-material cells. Time step~$\Delta t$ is chosen as~$T / 40$. This solution was found to be (almost) mesh independent and will serve as the  reference solution. The center cutline profiles~$p_*\big((x,0.5), t\big)$   are shown in the Figure~\ref{fig:transient:ref} (right) for several values of~$t \in (0, T]$.

For ASC(0), ASC(1), arithmetic and harmonic homogenization methods we use two Voronoi meshes with~$h = 1.5 \times 10^{-1}$ (124 base\,cells) and $h = 8.1 \times 10^{-2}$ (465 base\,cells).
The coarse mesh contains 20 multi-material cells and two macro-faces sharing three materials. The fine mesh contains 50 multi-material cells and no three material macro-faces. For all these methods we use 20 time steps, $\Delta t = T / 20$. Results are shown in Figures~\ref{fig:transient:comp}\,--\,\ref{fig:transient:err}.

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2/movie_frames/ref/frame_1s34.png}
		\caption{Reference}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/arithmetic_homo/frame_1s43.png}
		\caption{Arithmetic homo.}	
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/harmonic_homo/frame_1s43.png}
		\caption{Harmonic homo.}	
	\end{subfigure}%
	\par
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/asc0/frame_1s43.png}
		\caption{ASC(0)}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/asc1/frame_1s43.png}
		\caption{ASC(1)}	
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.3\linewidth}
		\centering
		\includegraphicsw{transient2-1/movie_frames/asc_diff/frame_1s43.png}
		\caption{ASC(0,\,1) difference}	
	\end{subfigure}%
	\caption{Comparison of the discrete solutions~$p_h$ for~$h = 1.5 \times 10^{-1}$, $t = 1.25$ \label{fig:transient:comp}}
\end{figure}
\begin{figure}[h]
 	\centering
 	\vskip 1cm
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2-1/movie_slices_frames/frame_0s25.png}
 		\caption{Coarse mesh, $t = 0.25$}
 	\end{subfigure}%
 	\hfill
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2-1/movie_slices_frames/frame_1s50.png}
 		\caption{Coarse mesh, $t = 1.50$}	
 	\end{subfigure}%
 	\par
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2/movie_slices_frames/frame_0s25.png}
 		\caption{Fine mesh, $t = 0.25$}
 	\end{subfigure}%
 	\hfill
 	\begin{subfigure}{.45\linewidth}
 		\centering
 		\includegraphicsw{transient2/movie_slices_frames/frame_1s50.png}
		\caption{Fine mesh, $t = 1.50$}	
	\end{subfigure}%
	\caption{Comparison of the cuts~$p_h\big((x,0.5), t\big)$ of discrete solutions for coarse and fine meshes, $h = 1.5 \times 10^{-1}$ and $h = 8.1 \times 10^{-2}$ 		\label{fig:transient:cuts}}
\end{figure}
